---
title: "R Tip of the Day: visdat"
subtitle: "Visualizing Data Quality & Missingness"
author: "Romina Amirian"
date: "April 2026"
format:
  revealjs:
    slide-number: true
execute:
  echo: true
---

## 

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(visdat)
library(palmerpenguins)
```

##  {auto-animate="true"}

::: {style="text-align: center; margin-top: 150px;"}
# visdat

## Visualizing Data Quality in R
:::

## What is visdat?

**visdat** = Visualize + Data

It helps you see what is inside your dataframe at a glance

. . .

**Without visdat:**

-   `str()` gives numbers

-   `summary()` gives numbers

-   Hard to see problems

. . .

**With visdat:**

-   One picture = instant understanding

-   Like an X-ray for your data

## Think of it this way

Would you rather read a spreadsheet of missing values?

. . .

Or look at a heatmap?

. . .

visdat gives you the heatmap.

## Meet the penguins data

```{r}
penguins
```

## vis_dat() - The X-Ray

```{r}
vis_dat(penguins)
```

## vis_miss() - Finding Missing Data

```{r}
print(vis_miss(penguins, sort_miss = TRUE))

```

## vis_guess() - Catching Hidden Errors

```{r}
problem_penguins <- penguins |>
  slice(1:100) |>
  mutate(
    bill_length_mm = as.character(bill_length_mm),
    bill_length_mm = if_else(row_number() == 42, "missing", bill_length_mm)
  )

vis_guess(problem_penguins)
```

## vis_cor() - Correlation Heatmap

```{r}
penguins |>
  select(where(is.numeric)) |>
  vis_cor()
```
